---
title: "How to use optionsLibrary"
author: "Tom Smit"
date: "2025-11-28"
output: html_document
vignette: >
  %\VignetteIndexEntry{Option Pricing Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\

# Introduction

This vignette explains how to use the `optionsLibrary` package to price European options and compute Greeks, including a simple application.\

The main syntax is defined as follows:

- Spot price $S_0$ = S

- Strike price $K$ = K

- Risk-free rate $r$ = r

- Dividend yield $\delta$ = q

- Volatility $\sigma$ = sigma

- Time to maturity $T$ = T

\

Now, load `optionsLibrary`.

```{r}
library(optionsLibrary)
```

\

# Black-Scholes pricing
Computing a European call:
```{r}
S <- 100
K <- 100
r <- 0.03
q <- 0
sigma <- 0.2
T <- 0.5
```

```{r}
call_price <- bs_price(S, K, r, q, sigma, T, 'call')
call_price
```
\

Same for a European put:
```{r}
put_price <- bs_price(S, K, r, q, sigma, T, 'put')
put_price
```

\

# Option Greeks
Greeks are easily computed in a similar fashion:
```{r}
call_greeks <- bs_greeks(S, K, r, q, sigma, T, 'call')
call_greeks
```

```{r}
put_greeks <- bs_greeks(S, K, r, q, sigma, T, 'put')
put_greeks
```
\
Output is in list form, allowing to isolate a specific greek:
```{r}
call_greeks$Delta
```

\

# Binomial tree pricing

```{r}
n <- 100
```

```{r}
call_binomial <- binomial_price(S, K, r, q, sigma, T, n, 'call')
call_binomial
```

```{r}
put_binomial <- binomial_price(S, K, r, q, sigma, T, n, 'put')
put_binomial
```
\

# Binomial convergence to Black-Scholes
An easy application Using `optionsLibrary` is to plot the convergence of the binomial price to the Black-Scholes price:
```{r}
n_steps <- seq(5, 500, by = 5)
binomial_prices <- numeric(length(n_steps))
blackscholes_price <- bs_price(S, K, r, q, sigma, T, 'call')

for (i in seq_along(n_steps)) {
  n <- n_steps[i]
  binomial_prices[i] <- binomial_price(S, K, r, q, sigma, T, n, 'call')
}
```

```{r, fig.cap='Binomial convergence', fig.alt='Line plot showing binomial price converging to Black-Scholes price'}
plot(n_steps, binomial_prices, type = 'l', col = '#001c3d', lwd = 2, main = 'Convergence of binomial price', xlab = 'Number of steps (n)', ylab = 'Option price')
abline(h = blackscholes_price, col = '#e84e10', lwd = 1.5)
legend('topright', legend = c('Binomial price', 'Black-Scholes price'), col = c('#001c3d', '#e84e10'), lty = c(1, 1), lwd = c(3, 3))
```

